Asymmetric relationships and their effects on coexistence

### 2023-07-03 paper/RefereeResponses1.tex

\noindent C18) L49: This is not immediately clear to me why (obviously, you’ve shown this in the other paper). Is it possible to briefly state why? \\

\noindent ***\emph{We spent 15-20 minutes or so trying to add an explanation that was succinct enough to not break
the flow of the paper while still being informative, but we could not do it. Any more than a short sentence would break the 
flow, and we felt that all the one-sentence explanations we could come up with were unclear. 
FYI, in a nutshell the idea is as follows. Assume that low values of an environmental variable reduce the 
growth rates of local populations and high values elevate those growth rates. In that case one may
expect that left-tail associations increase metapopulation extinction risks because then environmental
``catastrophes'' are spatially synchronized, presumably reducing the potential for rescue effects. Whereas
``bonanzas'' (extremely good conditions, with a lot of growth) are much less correlated. The case where
high values of the variable reduce the growth and low values elevate it is similar, but it's right-tail association
that elevates extinction risk. This is most of the explanation, but it gets a bit more nuanced, still - 
see the paper for additional details. Thanks for your interest, by the way.} \\

\noindent C19) L61: What if not? Or many readers have only rough understanding of it, maybe quickly refresh or refer to later \\

\noindent ***\emph{It is not necessary for the reader to understand what these are at this point, and,
in fact, the reader does not need to understand what relative nonlinearity is at all for the main text
and main ideas of the paper. We apologize for giving the false impression, here, that more background knowledge was needed
than is actually the case. We eliminated the reference to relative nonlinearity and indicated to the 
reader that storage effects would be conceptually reviewed below (a couple paragraphs down) and then 
defined oihn Theory, which they are.} \\

@@

\noindent C20) Sidenote: After writing these few comments it appears to me that you seem to have a very knowledgeable reader in mind, which may not be the case for a general journal as Ecology letters. \\

\noindent ***\emph{Well, we appreciate the referee's effort to spur us to produce the clearest 
paper possible, but, of the comments 
C16-C19 above, only C19 relates at all to whether this paper is written for a general ecological audience,
and we have revised the text C19 mentions so as to robustly address that comment (see response above). 
The manuscript text mentioned in C17-C18 was just linking this paper to other prior work on ATAs. 
It's not necessary for the reader to understand that work to understand this paper, the point is just to make 
the link to that other literature, so that readers can explore it if they so choose, 
and to show that ATAs have been found to be important in ecology, generally, to motivate the study.
Furthermore, we don't think our paper is any more specialized than the papers Ellner et al (2016, 2019) which
we cite, and which inspired our paper, and which were published in Ecology Letters. Nor is it more 
specialized than many other papers we have read in Ecology Letters. Quantitive approaches are absolutely 
central in ecology these days, and increasingly so. It makes no sense to us to sideline them into 
"specialist journals", which few people read, especially when they are about mainstream topics such
as coexistence. Nevertheless, we have gone through the paper and redoubled our efforts to make the 
paper broadly accessible.} \\
%We have to go through the paper and redoubled our efforts to make the paper broadly accessible.

\noindent C21) L64: Narwani 2013 is actually not a good reference because it did not include any temporal variation, a better one would be Letten et al. 2019 in PNAS. \\

\noindent ***\emph{We replaced the reference.} \\
%Jasmin can you please do this?

\noindent C22) L74: The optimum might actually be the same, but they need different responses. \\

\noindent ***\emph{Thanks for pointing this out, we have corrected our language.} \\

\noindent C23) Eq. 1: I would not use any equations in the introduction, and in general this part of the introduction did not feel like an introduction at all and might require a different sub-section \\

\noindent ***\emph{We disagree, and would prefer not to make this change, for several reasons. 
First, the lottery model is a very simple model, at the same level of complexity
as models introduced in essentially all introductory 
undergraduate ecology courses (e.g., the Lotka-Volterra competition model), so there is no argument that 
the readership of Ecology Letters cannot easily understand the material presented here.
Second, this material serves a clear motivating purpose, most appropriately placed in the Introduction. 
The last author (DCR) has found, from decades of pitching 
quantitative papers to general ecological audiences, that clear motivation is essential - it is crucial to demonstrate 
clearly that an effect one plans to elaborate is present, and potentially important, prior to asking the reader to engage 
the kind of mental energy necessary to read the guts of the paper. So this 
paragraph is for showing that the effect of ATAs on coexistence can be real, at least for a simple 
example. That goal is convincingly 
accomplished by the paragraph. Given that the math is simple, and also more precise and definitive than 
qualitative verbal arguments that usually take up space in Introductions, we think this material is actually
more beneficial to the readership and more effective for the intended motivational purpose than would be more words
(and we see based on C25 below that the referee appears to agree with us here). 
Third, quantitative approaches are central in ecology these days, and increasingly so. The automatic tendency
to want to exclude mathematics of any kind from the Introduction, especially when the math is simple,
accessible, and serves a clear motivational purpose, seems like a dated bias to us. 
Finally, we firmly believe that ecology as a discipline gets stronger as it continues the clear
trend of the last few decades of incorporating more quantitative approaches. We believe we
benefit the field by challenging assumptions and biases such as categorically not putting equations 
in the Introduction.
Overall, we think making this suggested change would make the paper weaker, and would pander to an
outmoded and unhelpful tradition. So we will not make the change unless the 
editor insists on it as a condition of publication.} \\

\noindent C24) L93: I suggest ``Note that $N = N_1(t) + N_2(T)$ is constant through time''. \\

\noindent ***\emph{OK, we made this change, even though the same equation already appears just a few lines
above this spot. Perhaps the referee is right that the reminder will benefit some readers.} \\

\noindent C25) L95-99: I think this is a great teaser to get the readers to later dive in into your mathematics \\

\noindent ***\emph{Thank you! This is exactly our purpose, here, and we are pleased it has been effective.
Relating back to our response to C23, we think this purpose cannot really be convincingly accomplished
without first stating the lottery model, since the result uses that model. } \\

@@

\noindent C) L126: Do you do this in the appendix? \\

\noindent ***\emph{No. We can see how it would be done, but we have not formalized the mathematics. 
To avoid making claims that we have not fully supported
with rigorous mathematics or other formal methods, we rephrased to say the iid assumption is 
``probably not necessary''. Thanks for spurring us to do this.} \\

@@

\noindent C) Eq 2.: I thought we also have to remove the effect of C varying alone and E varying alone? Is this included in the $\#$ term automatically? \\

\noindent ***\emph{No, the equation is correct as is. Please see equation 10 and related equations of 
Ellner et al, 2019, Ecology Letters. Yes, the effects of $C$ and $E$ varying are included in 
the $\#$ term because that terms involves both of them varying (though independently).} \\

@@

\noindent C) Eq. 6: Maybe highlight that you use different brackets ([EC] versus (EC)), or probably even better, make 5 and 6 one equation using curly under brackets. \\

\noindent ***\emph{Yes, it is a good idea to make it completely clear to the reader that these are 
used distinctly. We now do so, though we do it above this point in the text - it is better, in our view,
to do this the first time we use the hard brackets. So thanks for pointing this out.
We do not know what the referee means by ``using curly under brackets'' so have not done anything
in response to that part of this comment.} \\

@@

\noindent C) L230 $\mu_1-\mu_2$ is negative, why not take $\mu_2-\mu_1$? \\

\noindent ***\emph{Well, we can do this if the editor insists, but to us, avoiding the use of
negative numbers is insufficient reason to have to undertake the substantial and time-consuming 
work (and potential for error) 
of re-doing several figures, and revising many formulas (in the main text and SI) which currently use the other convention.
After all, negative numbers are basic, and represent no real impediment to the average ecologist understanding the 
paper. And we could instead spend that time working on the concepts underlying the next paper.} \\

@@

\noindent C) L316: So why do you show values of $\sigma >6$? And also maybe highlight this in the figure Figure 1d and e: I’m semi convinced by these distributions showing ATA. In 1d it simply appears that the variance scales with the density. 1e is quite unconvincing to me. That’s generally fine, but figure 1 seems to be a conceptual figure, so I wonder whether you don’t have a better example? \\

\noindent ***\emph{Regarding the first referee question listed above, well, what we are 
trying to do here is accurately portray our results about
when ATAs were important and when they were not. To us, the most forthright way of doing that is to
show \emph{all} the results, including the values of $\sigma$ which are probably unrealistic, and mention
(as we do) that some values are unrealistic. The point of this paragraph is to outline the circumstances
we have explored, using our lottery-model approaches, for which ATAs were important or not, and including all
our results seems important when such is our goal. Furthermore, there's no real cost to showing
all the $\sigma$'s we explored, it does not take up more manuscript space to do so.
Some readers may also take the lesson, for future work, from these explorations that one of the things that probably 
matters for continuing to explore when ATAs are important and when not is probably the strength
of the noise impacting the system. Finally, whether a value of $\sigma$ is ``unrealistic'' is 
a bit of judgement call rather than a formal cutoff, to we hesitate to censor results 
based on such a thing.

Regarding the second part of the referee comment, pertaining to Figure 1d and e, it is perplexing to us that 
this is mentioned here,
as the paragraph at L316 is not directly related to Fig. 1d-e. That figure
shows motivational empirical examples, distinct from the lottery model work discussed around L316. Nevertheless, we take
seriously the referee's comment that they are are not convinced there are ATAs in Fig. 1d-e, and ...
%Jasmin, let's discuss what to do here. We could work back to what specific result this is in 
%the original paper and extract from p-value info, or we could come up with a better example,
%or something else, not sure, let's discuss.
} \\

@@

\noindent C) Figure 2: I recommend to add a line at $y = 0$. Also the different subgraphs have different $y$-scales which I’m not sure is beneficial, but unsure myself. \\

%Jasmin, isn't there already such a line? It just gets overplotted sometimes, right? But you can see it on
%panel (e), for instance, right? I've written a response assuming that, but let's briefly discuss.
\noindent ***\emph{There is already a line at $y=0$, you can see it, for instance, on panel (e). 
It gets unavoidably overplotted on panels for which one of the plotted quantities is close to 0. The reason
for different $y$-axis extents is so that you can see the variation within each panel with maximum 
clarity. The main point, here, is to compare the relative contributions of different components
to coexistence, so it's important to show within-plot variation (between lines and across 
values of the $x$-axis) clearly. We thank the referee for expressing their own uncertainty, which was
useful to be aware of.} \\

@@

\noindent C) Figure 2 and 3: It appears that ATA is almost never the relevant factor, however I wonder whether this is because ATA is relatively weak, or whether the same analysis of say $\delta_i$ would lead to a similar pattern, it might be worth creating these graphs in the appendix to strengthen your case. \\

\noindent ***\emph{For Fig. 2, the referee is right that ATAs are only rarely important, as explained in the text 
(see first paragraph of section 4.1). However, for Fig 3, the referee is incorrect. In that figure, the ATA term
is usually one of the largest-magnitude terms. Our results as to the importance of ATAs for coexistence 
contrast between our log-normal and beta funcundities lottery models, as explained in the last paragraph
of section 4.1.} \\

@@

\noindent C) Figure 2: ATA is always positive, so how can it lead to ATA exclusion? (or ATA rescue in case I misunderstood). \\

\noindent ***\emph{The referee appears to have misinterpreted the figure. The ATA terms is, in fact, usually
negative for this figure. The colors of lines corresponding to different terms are in the legend, and, in the caption,
it is reviewed which of these corresponds to the ATA term. Since these items are clearly labelled, and since 
the $y=0$ line is also already present, we are unsure how to reduce the chances for other readers to make
such errors. But we're happy to hear specific suggestions.} \\

### 2023-07-03 paper/Paper.Rnw
continued working on responses

%Intro to competition, coexistence, MCT, etc.
Competition, coexistence, and the maintenance of biodiversity are research topics at the core of modern ecology 
\citep{MacArthur_1958, Hutchinson_1961, Schmida_1984}. The competitive exclusion principle states 
that two species competing for the same limiting resource cannot coexist at constant densities - one will 
instead outcompete the other and drive it extinct \citep{Hardin_1960}.
Modern coexistence theory [MCT; \cite{Chesson_1994,Chesson_2000,Barabas_2018}] is a 
framework for understanding how species nevertheless coexist, and biodiversity is maintained, in real ecosystems,
through the niche partitioning effects of environmental variation.
MCT breaks down coexistence into the contributions of various mechanisms, with names such as ``storage effects'' \citep{Chesson_1994} which will be familiar to some readers (storage effects are also defined below). MCT has been
applied to several systems, including annual \citep{Angert_2009, Hallett_2019} and perennial plants \citep{Adler_2006, Adler_2010}, 
tropical trees \citep{Usinowicz_2012}, plankton \citep{Caceres_1997, Descamps_2005, Narawani_2013}, and corals \citep{Alvarez-Noriega_2020}. 
Although MCT is somewhat mathematically complex, recent extensions make it more accessible \citep{Ellner_2016,Ellner_2019}.

%Intro to storage effects and why they might be influenced by ATAs
It is possible to see, intuitively, 
why ATAs may influence one mechanism of coexistence in particular, storage effects. We review, conceptually, 
what storage effects are \citep{Chesson_1994,Chesson_2012,Ellner_2016,Barabas_2018}, and explain why ATAs 
may be relevant to them. Storage effects allow competing species to fluctuate in abundance, without any going extinct,
by providing a relative benefit to whichever species is currently rare \citep{Chesson_2000}.
Storage effects hinge on positive covariation between the fluctuating quality of the environment
for a species (E) and the strength of competition (C) experienced by that species (called EC covariance);
and on the assumption that optimal environmental conditions for distinct species are also distinct,
so that no two species find the same conditions to be optimal. Then, good environmental
conditions for a currently common species are paired with strong intraspecific competition, because of
positive EC covariance and the commonness of the species. But good environmental conditions for 
a currently rare species are not as closely accompanied by strong competition because the rarity of
the species limits competitive influence. The resulting asymmetry is what provides the relative benefit to the
rare species. Storage effects can be quantified by measuring to what extent EC covariance contributes to 
differences between currently rare and currently common species growth rates \citep{Chesson_1994,Ellner_2016}.  
ATAs may, in principle, influence storage effects simply because storage effects hinge on EC covariation, 
and ATAs, when they occur between environment and competition variables, can be an aspect of EC covariation.

%Lottery model and results based on it
To further demonstrate the plausibility of ATAs altering coexistence between species, we carried out 
simple simulations using a two-species model, the well-known lottery model. The simulations described in this
paragraph are sufficient to demonstrate that ATAs influence coexistence in some manner; the
precise nature of that influence is the subject of the rest of the paper.
The lottery model, which was originally introduced as a model of coral reef fish competing for a set of $N$ breeding territories, 
is as follows.
Letting $N_i(t)$ denote the adult population density of species $i=1,2$ at time $t$, and defining $N = N_1(t)+N_2(t)$,
model equations are
\begin{equation}
N_i(t+1)=(1-\delta)N_i(t)+\delta N \frac{B_i(t)N_i(t)}{B_1(t)N_1(t)+B_2(t)N_2(t)}  \label{model_eq}
\end{equation}
for $i=1,2$. Here, $\delta$ is an adult mortality rate, and $B_i(t)$ is the fecundity of species $i$ at time $t$. 
The model postulates that adults die at rate $\delta$ at each time step, and are replaced by juveniles 
in proportion to the reproductive outputs of the two species that year. Note that $N$ is constant through time. 
Notation for the whole paper is summarized in Table \ref{tab:notation}.
We assume for simplicity that the random variables $(B_1(t),B_2(t))$ are independent and identically 
distributed (iid) through time. In three distinct simulations (Fig. \ref{fig:pedagog1}f-h), 
$B_1$ and $B_2$ were, respectively, 
left- (Fig. \ref{fig:pedagog1}a), right- (Fig. \ref{fig:pedagog1}c) and symmetrically (Fig. \ref{fig:pedagog1}b)
tail associated, while the same marginal distributions for these fecundities were used in all 
simulations (see Methods
for details). Thus, the model setup was identical in the three simulations except for different ATAs. 
ATAs substantially influenced aspects of species coexistence in these simulations (Fig. \ref{fig:pedagog1}f-h).

@@

We develop our new theory for a general two-species model, later applying it to specific scenarios.
Our mathematical notation parallels
that of \cite{Ellner_2016,Ellner_2019}. It is assumed that the growth rate $r_i(t)=\ln(N_i(t+1)/N_i(t))$ 
of species $i$ can be written as an increasing function of
an environment-dependent factor $E_i(t)$ and as a decreasing function of a quantity representing
competitive pressure on species $i$, $C_i(t)$, i.e., $r_i(E_i,C_i)$. We also assume, for simplicity, 
that the $E_i(t)$ are iid through time, and likewise for the
$C_i(t)$. Though the iid assumption is not necessary, it simplifies the
presentation of new ideas.

@@

We now specify how to decompose storage effects into
contributions due to ATAs and due
to $EC$ correlation \emph{per se}. 
Storage effects for species $i$ were computed by comparing
differences between rare and common species mean growth rates for the original model and for a 
modified model for which the components of $(E_i,C_{i \bs i})$ and those of $(E_j,C_{j \bs i})$
were rendered independent of each other [see (\ref{eq:storage_effects_gen})]. 
Because these components were rendered 
completely independent in the modified model, storage effects quantify the contribution of the 
whole $EC$ relationship, in all its aspects, to the rare/common species growth rate
difference. But we can instead consider distributions $(E_i^\nat,C_{i \bs i}^\nat)$ and 
$(E_j^\nat,C_{j \bs i}^\nat)$ with the same
marginal distributions, respectively, as $(E_i,C_{i \bs i})$ and $(E_j,C_{j \bs i})$,
and with some but not all aspects of the relationship 
between the components altered relative to the original distributions. Specifically, asymmetries of 
tail association are removed -- see SI section \ref{SI-sect:remove_asymmetries} for how this is done. We then define 
$\olrn_{i \bs i}$ and $\olrn_{j \bs i}$ based on a modified model using  $(E_i^\nat,C_{i \bs i}^\nat)$
and $(E_j^\nat,C_{j \bs i}^\nat)$, and we consider the decomposition 
\begin{align}
\Delta_i^{(EC)} &= [(\olr_{i\bs i} - q_{ij} \olr_{j \bs i})-(\olrn_{i \bs i}-q_{ij} \olrn_{j \bs i})]+
[(\olrn_{i \bs i}-q_{ij} \olrn_{j \bs i})-(\olrs_{i \bs i}-q_{ij} \olrs_{j \bs i})] \label{eq:storage_effects_further1}\\
&= \Delta_i^{[EC]} + \Delta_i^{[E \nat C]}.\label{eq:storage_effects_further}
\end{align}
The term $\Delta_i^{[EC]}$ is the component of storage effects due to 
ATAs between $E_i$ and $C_{i \bs i}$, and between $E_j$ and $C_{j \bs i}$. We will refer to this term as 
the \emph{ATA effect on coexistence}. 
The term $\Delta_i^{[E \nat C]}$ is the component of storage effects due to $EC$ 
correlation \emph{per se}, remaining after having removed the effects of ATAs.
The notation $\nat$ was chosen because this new symbol is part of the symbol $\shp$ (i.e., two of its four lines), 
corresponding to the removal of part of the relationship between $E$ and $C$ components (namely, ATAs).
The new symbol should correspondingly be pronounced ``partial sharp''.
The brackets in $\Delta_i^{[E \nat C]}$ and $\Delta_i^{[EC]}$ are intended to indicate that 
(\ref{eq:storage_effects_further}) is a decomposition within a decomposition, i.e., (\ref{eq:storage_effects_further}) is a decomposition
of storage effects, which are themselves a term in the MCT decomposition \citep{Ellner_2019} -- see 
below for additional details.

@@

The authors thank Shyamolina Ghosh for suggesting the research problem, and Lawrence Sheppard, Max Castorani,
Jon Walter, Jude Kastens, Vadim Karatayev, Emily Arsenault, Jordan Rodriguez, Lauren Hallett, and 
Robin Snyder for helpful discussions. 
The authors especially thank Steve Ellner for key advice at several watershed moments in the development of the project. 
The authors were partly supported by US National
Science Foundation grants 1714195 and 2023474, the James S McDonnell Foundation, the California Department of Fish and
Wildlife Delta Science Program, and the Alexander von Humboldt Foundation.

>>>>

%Intro to competition, coexistence, MCT, etc.
Competition, coexistence, and the maintenance of biodiversity are research topics at the core of modern ecology 
\citep{MacArthur_1958, Hutchinson_1961, Schmida_1984}. The competitive exclusion principle states 
that two species competing for the same limiting resource cannot coexist at constant densities - one will 
instead outcompete the other and drive it extinct \citep{Hardin_1960}.
Modern coexistence theory [MCT; \cite{Chesson_1994,Chesson_2000,Barabas_2018}] is a 
framework for understanding how species nevertheless coexist, and biodiversity is maintained, in real ecosystems,
through the niche partitioning effects of environmental variation.
MCT breaks down coexistence into the contributions of various mechanisms, with names such as ``storage effects'' \citep{Chesson_1994} which will be familiar to some readers (storage effects are also reviewed conceptually below and 
defined formally in Theory). MCT has been
applied to several systems, including annual \citep{Angert_2009, Hallett_2019} and perennial plants \citep{Adler_2006, Adler_2010}, 
tropical trees \citep{Usinowicz_2012}, plankton \citep{Caceres_1997, Descamps_2005, Narawani_2013}, and corals \citep{Alvarez-Noriega_2020}. 
Although MCT is somewhat mathematically complex, recent extensions make it more accessible \citep{Ellner_2016,Ellner_2019}.

%Intro to storage effects and why they might be influenced by ATAs
It is possible to see, intuitively, 
why ATAs may influence one mechanism of coexistence in particular, storage effects. We review, conceptually, 
what storage effects are \citep{Chesson_1994,Chesson_2012,Ellner_2016,Barabas_2018}, and explain why ATAs 
may be relevant to them. Storage effects allow competing species to fluctuate in abundance, without any going extinct,
by providing a relative benefit to whichever species is currently rare \citep{Chesson_2000}.
Storage effects hinge on positive covariation between the fluctuating quality of the environment
for a species (E) and the strength of competition (C) experienced by that species (called EC covariance);
and on the assumption that environmental responses for distinct species are also distinct,
so that no two species respond in the same way to environmental variation. Then, good environmental
conditions for a currently common species are paired with strong intraspecific competition, because of
positive EC covariance and the commonness of the species. But good environmental conditions for 
a currently rare species are not as closely accompanied by strong competition because the rarity of
the species limits competitive influence. The resulting asymmetry is what provides the relative benefit to the
rare species. Storage effects can be quantified by measuring to what extent EC covariance contributes to 
differences between currently rare and currently common species growth rates \citep{Chesson_1994,Ellner_2016}.  
ATAs may, in principle, influence storage effects simply because storage effects hinge on EC covariation, 
and ATAs, when they occur between environment and competition variables, can be an aspect of EC covariation.

%Lottery model and results based on it
To further demonstrate the plausibility of ATAs altering coexistence between species, we carried out 
simple simulations using a two-species model, the well-known lottery model. The simulations described in this
paragraph are sufficient to demonstrate that ATAs influence coexistence in some manner; the
precise nature of that influence is the subject of the rest of the paper.
The lottery model, which was originally introduced as a model of coral reef fish competing for a set of $N$ breeding territories, 
is as follows.
Letting $N_i(t)$ denote the adult population density of species $i=1,2$ at time $t$, and defining $N = N_1(t)+N_2(t)$,
model equations are
\begin{equation}
N_i(t+1)=(1-\delta)N_i(t)+\delta N \frac{B_i(t)N_i(t)}{B_1(t)N_1(t)+B_2(t)N_2(t)}  \label{model_eq}
\end{equation}
for $i=1,2$. Here, $\delta$ is an adult mortality rate, and $B_i(t)$ is the fecundity of species $i$ at time $t$. 
The model postulates that adults die at rate $\delta$ at each time step, and are replaced by juveniles 
in proportion to the reproductive outputs of the two species that year. Note that $N = N_1(t)+N_2(t)$ 
is constant through time. 
Notation for the whole paper is summarized in Table \ref{tab:notation}.
We assume for simplicity that the random variables $(B_1(t),B_2(t))$ are independent and identically 
distributed (iid) through time. In three distinct simulations (Fig. \ref{fig:pedagog1}f-h), 
$B_1$ and $B_2$ were, respectively, 
left- (Fig. \ref{fig:pedagog1}a), right- (Fig. \ref{fig:pedagog1}c) and symmetrically (Fig. \ref{fig:pedagog1}b)
tail associated, while the same marginal distributions for these fecundities were used in all 
simulations (see Methods
for details). Thus, the model setup was identical in the three simulations except for different ATAs. 
ATAs substantially influenced aspects of species coexistence in these simulations (Fig. \ref{fig:pedagog1}f-h).

@@

We develop our new theory for a general two-species model, later applying it to specific scenarios.
Our mathematical notation parallels
that of \cite{Ellner_2016,Ellner_2019}. It is assumed that the growth rate $r_i(t)=\ln(N_i(t+1)/N_i(t))$ 
of species $i$ can be written as an increasing function of
an environment-dependent factor $E_i(t)$ and as a decreasing function of a quantity representing
competitive pressure on species $i$, $C_i(t)$, i.e., $r_i(E_i,C_i)$. We also assume, for simplicity, 
that the $E_i(t)$ are iid through time, and likewise for the
$C_i(t)$. Though the iid assumption is probably not necessary, it simplifies the
presentation of new ideas.

@@

We now specify how to decompose storage effects into
contributions due to ATAs and due
to $EC$ correlation \emph{per se}. 
Storage effects for species $i$ were computed by comparing
differences between rare and common species mean growth rates for the original model and for a 
modified model for which the components of $(E_i,C_{i \bs i})$ and those of $(E_j,C_{j \bs i})$
were rendered independent of each other [see (\ref{eq:storage_effects_gen})]. 
Because these components were rendered 
completely independent in the modified model, storage effects quantify the contribution of the 
whole $EC$ relationship, in all its aspects, to the rare/common species growth rate
difference. But we can instead consider distributions $(E_i^\nat,C_{i \bs i}^\nat)$ and 
$(E_j^\nat,C_{j \bs i}^\nat)$ with the same
marginal distributions, respectively, as $(E_i,C_{i \bs i})$ and $(E_j,C_{j \bs i})$,
and with some but not all aspects of the relationship 
between the components altered relative to the original distributions. Specifically, asymmetries of 
tail association are removed -- see SI section \ref{SI-sect:remove_asymmetries} for how this is done. We then define 
$\olrn_{i \bs i}$ and $\olrn_{j \bs i}$ based on a modified model using  $(E_i^\nat,C_{i \bs i}^\nat)$
and $(E_j^\nat,C_{j \bs i}^\nat)$, and we consider the decomposition 
\begin{align}
\Delta_i^{(EC)} &= [(\olr_{i\bs i} - q_{ij} \olr_{j \bs i})-(\olrn_{i \bs i}-q_{ij} \olrn_{j \bs i})]+
[(\olrn_{i \bs i}-q_{ij} \olrn_{j \bs i})-(\olrs_{i \bs i}-q_{ij} \olrs_{j \bs i})] \label{eq:storage_effects_further1}\\
&= \Delta_i^{[EC]} + \Delta_i^{[E \nat C]}.\label{eq:storage_effects_further}
\end{align}
The term $\Delta_i^{[EC]}$ is the component of storage effects due to 
ATAs between $E_i$ and $C_{i \bs i}$, and between $E_j$ and $C_{j \bs i}$. We will refer to this term as 
the \emph{ATA effect on coexistence}. 
The term $\Delta_i^{[E \nat C]}$ is the component of storage effects due to $EC$ 
correlation \emph{per se}, remaining after having removed the effects of ATAs.
Note that brackets, $[\cdot]$, and parentheses, $(\cdot)$, are used distinctly in our notation,
throughout the paper - they are not interchangeable.
The notation $\nat$ was chosen because this new symbol is part of the symbol $\shp$ (i.e., two of its four lines), 
corresponding to the removal of part of the relationship between $E$ and $C$ components (namely, ATAs).
The new symbol should correspondingly be pronounced ``partial sharp''.
The brackets in $\Delta_i^{[E \nat C]}$ and $\Delta_i^{[EC]}$ are intended to indicate that 
(\ref{eq:storage_effects_further}) is a decomposition within a decomposition, i.e., (\ref{eq:storage_effects_further}) is a decomposition
of storage effects, which are themselves a term in the MCT decomposition \citep{Ellner_2019} -- see 
below for additional details.

@@

The authors thank Shyamolina Ghosh for suggesting the research problem, and Lawrence Sheppard, Max Castorani,
Jon Walter, Jude Kastens, Vadim Karatayev, Emily Arsenault, Jordan Rodriguez, Lauren Hallett, 
Robin Snyder, and three anonymous referees and the editor for helpful feedback. 
The authors especially thank Steve Ellner for key advice at several watershed moments in the development of the project. 
The authors were partly supported by US National
Science Foundation grants 1714195 and 2023474, the James S McDonnell Foundation, the California Department of Fish and
Wildlife Delta Science Program, and the Alexander von Humboldt Foundation.
